### 2024-05-27 figures_src/fv_grid.tex
updated cgrid figure

\documentclass[tikz,border=3mm]{standalone}
%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsmath package provides various useful equation environments.
\usepackage{amsmath}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm}

\begin{document}
	\begin{tikzpicture}[bullet/.style={circle,fill=#1,inner sep=1.5pt}] 
    \draw[->] (-1,-1) -- (5.5,-1) node[right] {$x$};
    \draw[->] (-1,-1) -- (-1,5.5) node[above] {$y$};
	\draw (0,0) grid[xstep=4, ystep=4] (4,4);
	\draw (0,0) grid[xstep=4, ystep=4] (4,4);
	\draw[dashed] (0,0) grid[xstep=2, ystep=2] (4,4);
    %\draw[dashed] (0,0) grid[xstep=2, ystep=2] (4,4);
	\draw[dashed] (-1,-1) grid[xstep=2, ystep=2]  (5,0);
	\draw[dashed] (-1,-1) grid[xstep=2, ystep=2]  (0,5);
	\draw[dashed] (-1,4) grid[xstep=2, ystep=2]  (5,5);
	\draw[dashed] (4,-1) grid[xstep=2, ystep=2]  (5,5);

    \node[left,font=\large] at (-1.2,0) {$y_{j-\frac{1}{2}}$};
    \node[left,font=\large] at (-1.2,2) {$y_{j}$};
    \node[left,font=\large] at (-1.2,4) {$y_{j+\frac{1}{2}}$};
    \node[below,font=\large] at (0,-1.2) {$x_{i-\frac{1}{2}}$};
    \node[below,font=\large] at (2,-1.2) {$x_{i}$};
    \node[below,font=\large] at (4,-1.2) {$x_{i+\frac{1}{2}}$};

%cell centroids
\node[bullet=black,label={[below, left]:$\rho$$,\phi$}, minimum size=0.25cm] at (2,2) {};

%u points
\node[bullet=blue,label={[below, left]:$\mathfrak{u}$}, minimum size=0.25cm] at (0,2.0) {};
\node[bullet=blue,label={[below, left]:$\mathfrak{u}$}, minimum size=0.25cm] at (4,2.0) {};

% v points
\node[bullet=red,label={[below, left]:$\mathfrak{v}$}, minimum size=0.25cm] at (2.0,0.0) {};
\node[bullet=red,label={[below, left]:$\mathfrak{v}$}, minimum size=0.25cm] at (2.0,4.0) {};

%corner points
\node[rectangle,fill] at (4,4) {};
\node[rectangle,fill] at (0,4) {};
\node[rectangle,fill] at (4,0) {};
\node[rectangle,fill] at (0,0) {};
	\end{tikzpicture}
\end{document}